---
title: "R.Python"
author: "MLDL"
date: "1/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, I am an R enthusiest to the core. Although, admittedly, this was likely a result of finding the limits of Stata while in graduate school, my love for code has been around for a long while. 

Much of my work (unfortunately not all... I still have to wrangle pivot tables in Excel and develop the occasional Tableau dashboard) these days is directly in R. I wrangle data, produce reports, iterrate various predictive models, and mostly train and employ a series of ML models for pure predictive purposes in a fairly R centric workflow. 

Nevertheless, I'm alwaysing trying to find connections to engrain what I've learned in other areas with what I know and love to be within R. 

For this example, and because a large bulk of many data scientists job is to wrangle data, I'll be doing a fairly simple EDA example between R and Python using the `reticulate` package and rmarkdown. 

This tutorial, if you will, will hopefully provide all that you need to get this up and running on your machine. I am, however, assuming you have a fully functional python install (and R.... obviously). 

Let's get started. 

First, we'll need two packages.  {[tidyverse](https://www.tidyverse.org)}, which will employ our R code, and the {[reticulate](https://rstudio.github.io/reticulate/)} package. 

You may find reticulate on CRAN. Also, our data today will comes from the famous Ames Housing pricing data, which can be found in their {AmesHousing} package. Well be using the processed version of the data. 

```{r message = FALSE}
#Packages Needed
library(tidyverse)
library(reticulate)
library(AmesHousing)

# create a new environment 
# conda_create("r-reticulate")

# install janitor

# conda_list()
# use_condaenv(condaenv = "r-reticulate")
# use_virtualenv("/Users/morganldebusk-lane/miniconda3/envs/r-reticulate/", required = TRUE)
# pf <- import("pandas")
# py_discover_config()
# matplotlib <- import("matplotlib")
```

Next, we need to configure how R knows where to employ python. First, lets see what options we have with the `py_config()` function from the reticulate package. 

```{r}
py_config()
```

Understanding which python I typically use, I need to tell R to use the miniconda3 version. For my set up, I simply copied the file path provided in the above output into the `use_python` command. 



```{r, include = FALSE}
use_python("~/Users/morganldebusk-lane/miniconda3/bin/python")
```

```{r}
use_condaenv(condaenv = "r-reticulate")
```
As far as I'm aware, that is all that is required to get things up and running. Lets give things a try. 

Today's data set if from the fairly simple, yet popular AMES housing prices dataset. 


```{r}
ames_prices <- ames_raw %>%
  #we only need a handful of variables for this example
  select(`Lot Frontage`:`Lot Shape`)
```

Now that we have our data, one of the first things I always do, and simply because I dont like to think too much about variable names, I put them all to lowercase and snakecase. This saves me a ton of time later on when I'm trying to write out variables and dont have to remember their peculiar case issues. 

In R, this calls for the famous {janitor} package and their `clean_names()` function that makes all column names lowercase and snake case. It does a whole bunch of other things automatically... like rename columns that started with a number into one that starts with a letter, etc. It's super useful. 

```{r}
ames_prices_clean_names <- ames_prices %>% 
  janitor::clean_names()
```



```{r}
head(ames_prices_clean_names, 5)
```

Using the original object 'ames_prices' we'll use python to make all column names lowercase and snakecase. 

```{r}
py_ames <- r_to_py(ames_prices)
```

```{python}
import pandas
py_ames_prices = r.ames_prices
print(r.ames_prices)
```

```{python}
py_ames_prices.columns = py_ames_prices.columns.str.lower().str.replace(' ', '_')

print(py_ames_prices)
```

Now that we have that, lets do some fairly simple aggregations to get to know the data better. 

I'd like to know the average `lot_frontage` and the different types of `street`. 

```{r}
ames_prices_clean_names %>%
  summarize(avg.lot = mean(lot_area, na.rm = TRUE))
```

```{r}
ames_prices_clean_names %>%
  group_by(street) %>% 
  tally()
```

```{python}
py_ames_prices.lot_frontage.mean()
```

```{python}
py_ames_prices.mean()

```

